Light versus Dark in Strong Lensing Galaxies: Dark matter halos that are rounder than their stars

### 2014-12-22 shapes/paper.tex
Justin: small edits throughout. Claudio's questions answered/addressed.

\title[Light versus Dark in Strong Lensing Galaxies]{Light versus Dark in Strong  Lensing Galaxies: Dark matter halos are rounder than their stars}

@@

\begin{abstract}
We measure the shape and alignment of stars and dark matter in 11 strong
lensing galaxies and find that in all cases the dark matter halos are rounder
than the light distribution over the range $R_e < R < 5R_e$.  Averaging over
larger radii, the lenses become increasingly elliptical both in their dark
matter and stellar components, as defined using the ratio $s$ of the largest to
smallest eigenvalues of the 2D moment of inertia tensor of the projected
lensing mass.  While dark matter halos are never more elliptical than $s_{dm} =
1.15$, their stars can extend to $s_* > 1.4$ Three systems, in particular, have
very high stellar ellipticity ($s_* > 1.6$) and correspondingly high alignment
between light and dark. One of these -- B1608 -- is a known merging pair; we
suggest that the other two (B0712 and B2016) may also be recent post-merger
systems.  Galaxies with high dark matter ellipticity and weak external shear
show very strong alignment between light and dark; those with strong shear
($\gamma \simgt 0.1$) can be very highly misaligned. This is reassuring since
isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must
explain the origin of very round dark matter halos and highly misaligned
systems. Such misalignments also present a new challenge for alternative
gravity theories in which the light and dark must necessarily be highly
correlated.

@@

The ellipticity and shape of stars relative to their host dark matter halo encodes information both about our cosmological model ($\Lambda$CDM) and galaxy formation \citep[e.g.][]{1994ApJ...431..617D,2001ApJ...551..294I,2004ApJ...611L..73K,2007MNRAS.378...55M,2007arXiv0707.0737D,2012MNRAS.424L..16L,2014JPhG...41f3101R}. `Dark-matter-only' (DMO) simulations in $\Lambda$CDM predict dark matter halos that are triaxial \citep{1991ApJ...378..496D,1992ApJ...399..405W,1996ApJ...462..563N,2002ApJ...574..538J}, with mean `shape parameter' $\langle q \rangle = (b+c)/2a \sim 0.8$ (where $a > b > c$ are the long, intermediate and short axes of the figure; \citealt{2007MNRAS.378...55M}). This corresponds to a typically {\it prolate} halo. However, including `baryons' (stars and gas) in the models produces halos that are significantly rounder and -- at least for disc galaxies -- well-aligned with the light distribution \citep{1991ApJ...377..365K,1994ApJ...431..617D,2007arXiv0707.0737D}. Halo shapes and alignments also constrain alternative gravity models \citep{2001MNRAS.327..552M,2004ApJ...610L..97H,2005MNRAS.361..971R,2012PhRvD..86h3507F,2013MNRAS.434.2971D}. If the visible light is the only source of gravity in a galaxy, then we expect the light and mass distribution to be highly correlated; if dark matter is present, however, such correlations can, at least in principle, be broken.

Strong lensing provides a unique probe of the alignment and shape of the total mass distribution in galaxies \citep[e.g.][]{1986ApJ...310..568B,1992grle.book.....S,1998ApJ...509..561K,2000ApJ...543..131K,2006ApJ...649..599K,2007AJ....134..668A,2008MNRAS.383..857F,2010ApJ...724..511A,2012MNRAS.424..104L}. For `red and dead' ellipticals that are largely devoid of gas, their baryonic content can be mapped through stellar population synthesis modelling of their light distribution \citep[e.g.][]{2005ApJ...623L...5F,2006ApJ...640..662T,2008MNRAS.383..857F}. This opens up the possibility of directly comparing the light and mass in strong lensing systems \citep{1998ApJ...509..561K,2008MNRAS.383..857F,2009ApJ...690..670T,2012A&A...538A..99S}. Previous work in the literature has found that the light and mass are well-aligned (though a mis-match of up to 10$^\circ$ is not uncommon; e.g. \citealt{2012A&A...538A..99S}). However, results on the ellipticity of light and mass agree less well, with \citet{2012A&A...538A..99S} finding a strong correlation and \citet{1998ApJ...509..561K} and \citet{2008MNRAS.383..857F} finding none. It is difficult, however, to compare the results between these different studies because they use different lens modelling techniques; different definitions of ellipticity; and different radii over which the shapes and alignments are probed. Furthermore, none to date have applied their methodology to mock data to determine the robustness of the results.

@@

Recently, we introduced a new non-parametric lens tool, \Glass\ \citep{2014arXiv1401.7990C}. Applying this to a large suite of mock data, we showed that mass and light can only reliably be disentangled in strong lensing systems if: i) there are at least four images; and ii) time delay data are available and/or the stellar mass dominates the potential over the region of the images. {\bf (CB. i) is the case for all the lenses except Q0957, ii) however is only the case for a minority of the sample. This sentence creates therefore some tension that is not addressed however (I don't know whether we should rephrase it or add another sentence))} In this paper, we collate data of the above quality, compiling a sample of 11 strong lens galaxies. We apply \Glass\ to these lenses to non-parametrically measure the shape and alignment of the stars and {\it dark matter}  in these lensing galaxies, for the first time. This differs from previous works that have all compared the light distribution with the total mass, rather than the dark matter. Since the stars often dominate the central potential, the total mass naturally correlates with the light, potentially masking theoretically interesting results about the dark matter distribution. Our comparison between light and dark is made possible by the fact that \Glass\ uses the light distribution as a prior on the mass map, ensuring that the dark matter mass is always positive.

@@

\Glass{} is a new `non-parametric' gravitational lens modelling framework
\citep{2014arXiv1401.7990C} that is capable of generating a distribution of
mass models given a set of basic assumptions, or priors.  {\bf Hence, \Glass\ does not yield the most probable mass model. Rather, we can describe the distribution of all the allowed models. (CB: I'VE ADDED THIS SENTENCE AS I FELT SOMETHING LIKE THAT MIGHT BE NEEDED AT THIS POINT. IT PROBABLY NEEDS TO BE IMPROVED THOUGH)} In this context,
non-parametric means that  many more parameters than data constraints are used
such that system of equations describing the lensing mass distribution is
under-constrained.  \Glass{} shares some aspects with an earlier code
\PixeLens{} \citep{2004AJ....127.2604S,2008ApJ...679...17C}, but has a completely
different code base. Its key improvements upon \PixeLens{} which are relevant
for this paper include:

@@

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item A reconstructed stellar mass map as lower mass bounds in every pixel.
\item The local density gradient is always pointing within 50$^{\circ}$ to the center and the slope is at most 0 everywhere. In other words, the density profile peaks in the center.
\item The fixed cosmology $\Omega_{m} = 0.28$, $\Omega_{\Lambda} = 0.72$, $\Omega_{k} = 0$, $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1)$.
\end{enumerate}

@@

% \item[(viii)] External point masses to model satellite galaxies outside but
close to the Einstein radius.

@@

\subsection{Measuring the shape and alignment of a lens}\label{sec:shapemethod}
We employ the same shape definition as in \citep{2014arXiv1401.7990C}. The shape parameter $s$ is defined as the ratio of the two eigenvalues $\lambda_1$ and $\lambda_2$, i.e.
\begin{equation}\label{eq:shapeestimate}
    s \equiv \frac{\lambda_{1}}{\lambda_{2}}, \ \ \ \mathrm{where} \ \lambda_{1} \geq \lambda_{2},
\end{equation}
of the inertia tensor of a mass distribution $M(\boldsymbol{\theta})$
\begin{equation}\label{eq:inertiatensor}
    I = \begin{pmatrix} \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
                        -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x} \end{pmatrix}.
\end{equation}
We use these estimators to quantify the shape of the stellar mass distribution and the dark matter distribution. We estimate the latter by reconstructing the total mass distribution with a lower mass prior from the stellar mass distribution and then subtracting the latter off. The shapes of the distributions are computed for all the pixels up to the outermost lensed image and for the pixels within different multiples of the half-light radius $R_e$. $R_e$ is the Petrosian radius containing half of the flux of the total light distribution. As terms in Eq.~\ref{eq:inertiatensor} are weighted with $r^2$, the outer pixels tend to dominate the estimate of the averaged shape within the considered radius. The shape estimates do not necessarily need to be the same within different radii. E.g. there are effects as isophotal twist \citep[e.g.][and references therein]{1978ComAp...8...27B} causing differences in the shape estimates between the central and the outer parts of the system. {\bf (CB: THESE LAST SENTENCES PROBABLY NEED TO BE REPHRASED TOO)}

Throughout this paper, we consider only the masses in the pixels up to one pixel length past the outermost image. Furthermore, we generate 10000 mass models for each system we consider and describe the distribution of all the individual estimates for the dark matter shapes. {\bf (CB: MAYBE NEEDS REPHRASING, SEE ABOVE)}

\section{Data}\label{sec:data}
Stellar-mass maps are taken from \citep{2011ApJ...740...97L}.  In that work, the light of the galaxy images in all available HST bands was first fitted using using GALFIT \citep{2002AJ....124..266P}, and the resulting light maps were used to obtain stellar-mass maps using a synthesis of the stellar-population models of \citep{2003MNRAS.344.1000B}.  The population synthesis marginalised over the star-formation epoch and time scales and over the stellar metallicity.  The IMF was taken to be the log-normal form given by \citep{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the inferred shape of the stellar mass.

As it is shown in \citep{2014arXiv1401.7990C}, the recovery of the shape information of the lens is only good if there are at least four images. Therefore, we only consider the quad lenses and the twin double \textit{Q0957+561} out of the sample presented in \citep{2011ApJ...740...97L}. Thus, the sample consists in total of 11 lenses. Each individual system is presented below.

@@

\textit{MG0414+0534} (hereafter \textit{0414}) is a passively evolving early-type galaxy \citep{1999AJ....117.2034T}. A very close luminous satellite galaxy was found north-west of the lens \citep{1993AJ....105....1S}. It however is more likely a foreground object \citep{2011MNRAS.413L..86C}.

@@

\textit{RXJ0911+0551} (hereafter \textit{0911}) is an almost circular early-type galaxy \citep{2012A&A...538A..99S}. It has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. X-ray emission can be detected and an analysis of which yields a temperature of 2.3 keV. There is also a satellite galaxy to the north-west direction \citep{2000ApJ...544L..35K}.

\textit{Q0957+561} (hereafter \textit{0957}) is a cD galaxy lying close to the center of a cluster with a high sprial galaxy-fraction \citep[e.g.][]{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}. Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the center of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. These results are consistent with the center of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay \citep[e.g.][]{2012A&A...540A.132S}. They however also find a three-day lag between the g- and r-bands, a disagreement at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

\textit{PG1115+080} (hereafter \textit{1115}) is an early-type galaxy \citep{2005ApJ...626...51Y}. It has measured time delays. In this work use recent estimates of the time delays by \cite{2010MNRAS.406.2764T} that differ from previously found values by e.g. \cite{1997ApJ...489...21B}. The environment of the lens was analyzed thouroughly \citep{2006ApJ...641..169M,2011ApJ...726...84W}. It is part of a small group of 13 members. Also X-ray emission was detected from the corresponding group, and it yields a temperature of 0.8 keV \citep{2004ApJ...610..686G}.

@@

\textit{B1608+656} (hereafter \textit{1608}) consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment has been analyzed and group with 8 (9 if merging galaxies are counted individually) was found \citep{2006ApJ...642...30F}. No significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. The data seems to indicate four other groups along the line of sight.

\textit{MG2016+112} (hereafter \textit{2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster which consists of 69 probable galaxies \citep{2003MNRAS.344..337T}. The clusters shows a high density of galaxies close to the lens in a south-east direction.

\textit{B2045+265} (hereafter \textit{2045}) is probably an elliptical galaxy \citep{2007MNRAS.378..109M}. It was initially classified as a late-type Sa galaxy \citep{1999AJ....117..658F}, the velocity dispersion however seems too high. As the source redshift is rather low, a large lens mass is required. \cite{2007MNRAS.378..109M} therefore conclude that it is more likely an elliptical galaxy. To the west of the lens, a group was found at a similar redshift as the lens \citep{1999AJ....117..658F}. There is furthermore evidence for a dwarf satellite galaxy \citep{2007MNRAS.378..109M}. As the measurements are inconclusive, we choose to not include this satellite.

\textit{Q2237+030} (hereafter \textit{2237}) is a barred spiral \citep{1988AJ.....95.1331Y}, the only non-early type lens in this sample. With only a redshift of $z=0.04$, it is the closest lens of the sample. Due to the low redshift, the probed physical scales are small. Therefore, the bulge component of the galaxy dominates the lens, what makes the stellar population synthesis modelling possible. No further inquiries into the environment were made. It is

Further information on the sample can be found in \citep{2011ApJ...740...97L,2012A&A...538A..99S}. The most important quantities of each system are given in Table~\ref{tab:lensproperties}. Due to the different angular separations of the images, different scales of the lens galaxies are probed.

\begin{table}
  \begin{center}
    \begin{tabular}{l r r r r l}
      Lens    & \multicolumn{1}{c}{$z_{L}$} & \multicolumn{1}{c}{$z_{S}$} & \multicolumn{1}{c}{$\Delta\theta$ [kpc]} & \multicolumn{1}{c}{$R_{L}$/$R_{e}$} & Environment \\ \hline
      0047 & 0.485 & 3.60 & 12.82 & $1.45\pm0.04$ & G(9) \\
      0414 & 0.960 & 2.64 & 16.01 & $1.85\pm0.05$ & ... \\
      0712 & 0.410 & 1.34 & 6.82  & $1.15\pm0.03$ & ... \\
      0911 & 0.769 & 2.8  & 23.16 & $3.09\pm0.05$ & C \\
      0957 & 0.356 & 1.41 & 29.98 & $3.51\pm0.04$ & C \\
      1115 & 0.310 & 1.72 & 10.76 & $2.86\pm0.06$ & G(13) \\
      1422 & 0.337 & 3.62 & 6.02  & $4.49\pm0.06$ & G(17) \\
      1608 & 0.630 & 1.39 & 13.92 & $1.82\pm0.01$ & G(8) \\
      2016 & 1.010 & 3.3  & 26.22 & $6.12\pm0.14$ & C(69) \\
      2045 & 0.870 & 1.28 & 14.46 & $1.48\pm0.03$ & ... \\
      2237 & 0.039 & 1.7  & 1.40  & $0.89\pm0.01$ & ... \\
    \end{tabular}
    \caption[width=\linewidth]{The most relevant lens properties for this work are listed \citep[for an expanded version of this table see][]{2011ApJ...740...97L}. References can be found in \S\ref{sec:data}. The columns are from left to right, the lens redshift $z_L$, the redshift of the source $z_S$, the maximum angular separation between two images $\Delta\theta$, the ratio of the radius of the outermost image $R_L$ and the effective Petrosian half-light radius of the lens $R_e$, and information on the environment of the lens. C and G denote a known cluster resp. group environment. The number in the brackets is the number of confirmed members.}
    \label{tab:lensproperties}
  \end{center}
\end{table}

>>>>

\title[Light versus Dark in Strong Lensing Galaxies]{Light versus Dark in Strong Lensing Galaxies: Dark matter halos that are rounder than their stars}

@@

\begin{abstract}
We measure the shape and alignment of stars and dark matter in 11 strong
lensing galaxies, finding in all cases that the dark matter halos are rounder
than the light distribution over the range $R_e < R < 5R_e$.  Averaging over
larger radii, the lenses become increasingly elliptical both in their dark
matter and stellar components, as defined using the ratio $s$ of the largest to
smallest eigenvalues of the 2D moment of inertia tensor of the projected
lensing mass.  While dark matter halos are never more elliptical than $s_{dm} =
1.15$, their stars can extend to $s_* > 1.4$ Three systems, in particular, have
very high stellar ellipticity ($s_* > 1.6$) and correspondingly high alignment
between light and dark. One of these -- B1608 -- is a known merging pair; we
suggest that the other two (B0712 and B2016) may also be recent post-merger
systems.  Galaxies with high dark matter ellipticity and weak external shear
show very strong alignment between light and dark; those with strong shear
($\gamma \simgt 0.1$) can be very highly misaligned. This is reassuring since
isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must
explain the origin of both very round dark matter halos and highly misaligned
strong lensing systems. Such misalignments also present a new challenge for alternative
gravity theories in which the light and dark must necessarily be highly
correlated.

@@

The ellipticity and shape of stars relative to their host dark matter halo encodes information both about our cosmological model ($\Lambda$CDM) and galaxy formation \citep[e.g.][]{1994ApJ...431..617D,2001ApJ...551..294I,2004ApJ...611L..73K,2007MNRAS.378...55M,2007arXiv0707.0737D,2012MNRAS.424L..16L,2014JPhG...41f3101R}. `Dark-matter-only' (DMO) simulations in $\Lambda$CDM predict dark matter halos that are triaxial \citep{1991ApJ...378..496D,1992ApJ...399..405W,1996ApJ...462..563N,2002ApJ...574..538J}, with mean `shape parameter' $\langle q \rangle = (b+c)/2a \sim 0.8$ (where $a > b > c$ are the long, intermediate and short axes of the figure; \citealt{2007MNRAS.378...55M}). This corresponds to a typically {\it prolate} halo. However, including `baryons' (stars and gas) in the models produces halos that are significantly rounder and -- at least for disc galaxies -- well-aligned with the light distribution \citep{1991ApJ...377..365K,1994ApJ...431..617D,2007arXiv0707.0737D}. Halo shapes and alignments also constrain alternative gravity models \citep{2001MNRAS.327..552M,2004ApJ...610L..97H,2005MNRAS.361..971R,2012PhRvD..86h3507F,2013MNRAS.434.2971D}. If the visible light is the only source of gravity in a galaxy, then we expect the light and mass distribution to be highly correlated; if dark matter is present, however, such correlations can at least in principle be broken.

Strong lensing provides a unique probe of the alignment and shape of the total mass distribution in galaxies \citep[e.g.][]{1986ApJ...310..568B,1992grle.book.....S,1998ApJ...509..561K,2000ApJ...543..131K,2006ApJ...649..599K,2007AJ....134..668A,2008MNRAS.383..857F,2010ApJ...724..511A,2012MNRAS.424..104L}. For `red and dead' ellipticals that are largely devoid of gas, their baryonic content can be mapped through stellar population synthesis modelling of their light distribution alone \citep[e.g.][]{2005ApJ...623L...5F,2006ApJ...640..662T,2008MNRAS.383..857F}. Furthermore, such systems are dense enough to produce strong lensing, opening up the possibility of directly comparing the light and mass in these galaxies \citep{1998ApJ...509..561K,2008MNRAS.383..857F,2009ApJ...690..670T,2012A&A...538A..99S}. Previous work in the literature has found that the light and mass are well-aligned (though a mis-match of up to 10$^\circ$ is not uncommon; e.g. \citealt{2012A&A...538A..99S}). However, results on the ellipticity of light and mass agree less well, with \citet{2012A&A...538A..99S} finding a strong correlation and \citet{1998ApJ...509..561K} and \citet{2008MNRAS.383..857F} finding none. It is difficult, however, to compare the results between these different studies because they use different lens modelling techniques; different definitions of ellipticity; and different radii over which the shapes and alignments are probed. Furthermore, none to date have applied their methodology to mock data to determine the robustness of the results.

@@

Recently, we introduced a new non-parametric lens tool, \Glass\ \citep{2014arXiv1401.7990C}. Applying this to a large suite of mock data, we showed that mass and light can only reliably be disentangled in strong lensing systems if: i) there are at least four images; and ii) time delay data are available and/or the stellar mass contributes significantly to the potential. In this paper, we collate data of the above quality, compiling a sample of 11 strong lens galaxies. We apply \Glass\ to these lenses to non-parametrically measure the shape and alignment of the stars and {\it dark matter}  in these lensing galaxies, for the first time. This differs from previous works that have all compared the light distribution with the total mass, rather than the dark matter. Since the stars often dominate the central potential, the total mass naturally correlates with the light, potentially masking theoretically interesting results about the dark matter distribution. Our comparison between light and dark is made possible by the fact that \Glass\ uses the light distribution as a prior on the mass map, ensuring that the dark matter mass is always positive.

@@

\Glass{} is a new `non-parametric' gravitational lens modelling framework
\citep{2014arXiv1401.7990C} generates a distribution of
mass models given a set of basic assumptions, or priors. Hence, \Glass\ does not yield the most probable mass model. Rather, it calculates an ensemble of all allowed models given the data and assumed priors. In this context,
`non-parametric' really just mean that many more parameters than data constraints are used
such that system of equations describing the lensing mass distribution is
under-constrained. \Glass{} shares some aspects with an earlier code
\PixeLens{} \citep{2004AJ....127.2604S,2008ApJ...679...17C}, but has a completely
different code base. Its key improvements upon \PixeLens{} which are relevant
for this paper include:

@@

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item A reconstructed stellar mass map as lower mass bounds in every pixel.
\item The local density gradient is always pointing within 50$^{\circ}$ to the center and the slope is at most 0 everywhere. In other words, the density profile peaks in the center.
\item The fixed cosmology $\Omega_{m} = 0.28$, $\Omega_{\Lambda} = 0.72$, $\Omega_{k} = 0$, $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1]$.
\end{enumerate}

@@

% \item[(viii)] External point masses to model satellite galaxies outside but
% close to the Einstein radius.

@@

\subsection{Measuring the shape and alignment of a lens}\label{sec:shapemethod}
We employ the same shape definition as in \citep{2014arXiv1401.7990C}. The shape parameter $s$ is defined as the ratio of the two eigenvalues $\lambda_1$ and $\lambda_2$, i.e.
\begin{equation}\label{eq:shapeestimate}
    s \equiv \frac{\lambda_{1}}{\lambda_{2}}, \ \ \ \mathrm{where} \ \lambda_{1} \geq \lambda_{2},
\end{equation}
of the inertia tensor of a mass distribution $M(\boldsymbol{\theta})$
\begin{equation}\label{eq:inertiatensor}
    I = \begin{pmatrix} \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{y} & -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} \\
                        -\sum_\theta M(\boldsymbol{\theta})\theta_{x}\theta_{y} & \sum_\theta M(\boldsymbol{\theta})\theta^{2}_{x} \end{pmatrix}.
\end{equation}
We use these estimators to quantify the shape of the stellar mass distribution and the dark matter distribution. We estimate the latter by reconstructing the total mass distribution with a lower mass prior from the stellar mass distribution and then subtracting the latter off. The shapes of the distributions are computed for all the pixels up to the outermost lensed image and for the pixels within different multiples of the half-light radius $R_e$. $R_e$ is the Petrosian radius containing half of the flux of the total light distribution. As terms in Eq.~\ref{eq:inertiatensor} are weighted with $r^2$, the outer pixels tend to dominate the estimate of the averaged shape within the considered radius. The shape estimates do not necessarily need to be the same within different radii -- e.g. effects such as isophotal twist \citep[e.g.][and references therein]{1978ComAp...8...27B} can cause differences in the shape estimates between the central and the outer parts of the system.

Throughout this paper, we consider only the masses in the pixels up to one pixel length past the outermost image. We generate in each case an ensemble of 10,000 mass models.

\section{Data}\label{sec:data}
Stellar mass maps are taken from \citep{2011ApJ...740...97L}. In that work, the light of the galaxy images in all available HST bands was first fitted using using GALFIT \citep{2002AJ....124..266P}, and the resulting light maps were used to obtain stellar mass maps using a synthesis of the stellar-population models of \citep{2003MNRAS.344.1000B}. The population synthesis marginalised over the star formation epoch and time scales and over the stellar metallicity. The IMF was taken to be the log-normal form given by \citep{2003PASP..115..763C}; a more bottom-heavy IMF would change the mass normalisation \citep[cf.][]{2014ApJ...793...96S} but not the inferred shape of the stellar mass.

As shown in \citep{2014arXiv1401.7990C}, the recovery of the shape information of the lens is only good if there are at least four images. Therefore, we only consider the quad lenses and the twin double \textit{Q0957+561} out of the sample presented in \citep{2011ApJ...740...97L}. This leads to a sample of 11 lenses. We discuss each of these systems, next.

@@

\textit{MG0414+0534} (hereafter \textit{0414}) is a passively evolving early-type galaxy \citep{1999AJ....117.2034T}. A very close luminous satellite galaxy was found north-west of the lens \citep{1993AJ....105....1S}. It, however, is more likely a foreground object \citep{2011MNRAS.413L..86C}.

@@

\textit{RXJ0911+0551} (hereafter \textit{0911}) is an almost circular early-type galaxy \citep{2012A&A...538A..99S}. It has measured time delays \citep{2002ApJ...572L..11H}. It lies on the outskirts of a cluster \citep{2001ApJ...555....1M}. X-ray emission can be detected and an analysis of which yields a temperature of 2.3 keV. There is also a satellite galaxy to the northwest \citep{2000ApJ...544L..35K}.

\textit{Q0957+561} (hereafter \textit{0957}) is a cD galaxy lying close to the centre of a cluster with a high spiral galaxy-fraction \citep[e.g.][]{1992MNRAS.254P..27G,1994A&A...291..411A,1998ApJ...504..661C}. Due to the large image separation, large physical scales are probed. As listed in Table 3 of \cite{2000ApJ...542...74K}, the position angles to the centre of the cluster of earlier lens reconstructions range between 51.8 deg and 67.8 deg measured north-through-east. These results are consistent with the centre of the X-ray emission from the cluster \citep{1998ApJ...504..661C}. The lens has a measured time delay \citep[e.g.][]{2012A&A...540A.132S}. They however also find a three-day lag between the g- and r-bands, a disagreement at the 2$\sigma$-level. They argue that this effect can be accounted for by the presence of a substructure and chromatic dispersion. We find that the results do not change significantly for either estimate and choose therefore the g-band measurement.

\textit{PG1115+080} (hereafter \textit{1115}) is an early-type galaxy \citep{2005ApJ...626...51Y}. It has measured time delays. In this work use recent estimates of the time delays by \cite{2010MNRAS.406.2764T} that differ from previously found values by e.g. \cite{1997ApJ...489...21B}. The environment of the lens was analysed thoroughly \citep{2006ApJ...641..169M,2011ApJ...726...84W}. It is part of a small group of 13 members. Also, X-ray emission was detected from the corresponding group and it yields a temperature of 0.8 keV \citep{2004ApJ...610..686G}.

@@

\textit{B1608+656} (hereafter \textit{1608}) consists of two merging galaxies. The main galaxy is an early-type galaxy which is disrupted by a smaller, probably late-type galaxy \citep{2003ApJ...584..100S}. The system has measured time delays \citep{2002ApJ...581..823F}. The environment has been analysed and a group with 8 (9 if merging galaxies are counted individually) was found \citep{2006ApJ...642...30F}. No significant X-ray emission was detected from the surrounding group \citep{2005ApJ...625..633D}. The data seem to indicate four other groups along the line of sight.

\textit{MG2016+112} (hereafter \textit{2016}) is a giant elliptical galaxy \citep{1984Sci...223...46L,1986AJ.....91..991S}. It is the farthest lens we consider in this sample. It lies in a cluster that consists of 69 probable galaxies \citep{2003MNRAS.344..337T}. The cluster shows a high density of galaxies close to the lens in a south-east direction.

\textit{B2045+265} (hereafter \textit{2045}) is probably an elliptical galaxy \citep{2007MNRAS.378..109M}. It was initially classified as a late-type Sa galaxy \citep{1999AJ....117..658F}, however the velocity dispersion seems too high. As the source redshift is rather low, a large lens mass is required. \cite{2007MNRAS.378..109M} therefore conclude that it is more likely an elliptical galaxy. To the west of the lens, a group was found at a similar redshift as the lens \citep{1999AJ....117..658F}. There is furthermore evidence for a dwarf satellite galaxy \citep{2007MNRAS.378..109M}. As the measurements are inconclusive, we choose to not include this satellite explicitly in our models.

\textit{Q2237+030} (hereafter \textit{2237}) is a barred spiral \citep{1988AJ.....95.1331Y}, the only non early-type lens in this sample. With a redshift of just $z=0.04$, it is also the closest lens of the sample. Due to its low redshift, the probed physical scales are small, making the bulge component of the galaxy dominant. However, this makes our stellar population synthesis model assumptions (i.e. that the stars are a simple stellar population) likely valid.

Further information on the sample can be found in \citet{2011ApJ...740...97L} and \citet{2012A&A...538A..99S}. The most important quantities of each system are given in Table~\ref{tab:lensproperties}. Due to the different angular separations of the images and the different angular diameter distances, in each lens galaxy we necessarily probe slightly different scales.

\begin{table}
  \begin{center}
    \begin{tabular}{l r r r r l}
      Lens    & \multicolumn{1}{c}{$z_{L}$} & \multicolumn{1}{c}{$z_{S}$} & \multicolumn{1}{c}{$\Delta\theta$ [kpc]} & \multicolumn{1}{c}{$R_{L}$/$R_{e}$} & Environment \\ \hline
      0047 & 0.485 & 3.60 & 12.82 & $1.45\pm0.04$ & G(9) \\
      0414 & 0.960 & 2.64 & 16.01 & $1.85\pm0.05$ & ... \\
      0712 & 0.410 & 1.34 & 6.82  & $1.15\pm0.03$ & ... \\
      0911 & 0.769 & 2.8  & 23.16 & $3.09\pm0.05$ & C \\
      0957 & 0.356 & 1.41 & 29.98 & $3.51\pm0.04$ & C \\
      1115 & 0.310 & 1.72 & 10.76 & $2.86\pm0.06$ & G(13) \\
      1422 & 0.337 & 3.62 & 6.02  & $4.49\pm0.06$ & G(17) \\
      1608 & 0.630 & 1.39 & 13.92 & $1.82\pm0.01$ & G(8) \\
      2016 & 1.010 & 3.3  & 26.22 & $6.12\pm0.14$ & C(69) \\
      2045 & 0.870 & 1.28 & 14.46 & $1.48\pm0.03$ & ... \\
      2237 & 0.039 & 1.7  & 1.40  & $0.89\pm0.01$ & ... \\
    \end{tabular}
    \caption[width=\linewidth]{The most relevant lens properties for this work are listed \citep[for an expanded version of this table see][]{2011ApJ...740...97L}. References can be found in \S\ref{sec:data}. The columns are from left to right, the lens redshift $z_L$, the redshift of the source $z_S$, the maximum angular separation between two images $\Delta\theta$, the ratio of the radius of the outermost image $R_L$ and the effective Petrosian half-light radius of the lens $R_e$, and information on the environment of the lens. C and G denote a known cluster or group environment, respecitvely. The number in the brackets is the number of confirmed members.}
    \label{tab:lensproperties}
  \end{center}
\end{table}
